---
title: "count_detections"
author: "Alvaro"
date: "2019-03-26"
site: workflowr::wflow_site
output:
  workflowr::wflow_html
---

## Introduction

The purpose of this document is to analyze th number of significant associations arising from GWAS, PrediXcan and MultiXcan methods
on several complex traits, on application of GTEx v8 data.

```{r preliminaries, echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(cowplot))
suppressPackageStartupMessages(library(ggrepel))
suppressPackageStartupMessages(library(RSQLite))
suppressWarnings(source("code/ab/_helpers.R"))
```

```{r trait_data, echo=FALSE}
gwas_metadata <- "data/gwas_metadata.txt" %>% r_tsv_ %>% filter(Deflation == 0)
trait_key <- gwas_metadata %>% select(trait=Tag, abbreviation=new_abbreviation, color=color, category=Category)

palette_ <- trait_key %>% select(category, color) %>% unique %>% .$color
categories_ <- trait_key %>% select(category, color) %>% unique %>% .$category
names(palette_) <- categories_
```

```{r association_data, echo=FALSE}
counts <- "data/gwas_hits/count.txt" %>% r_tsv_
counts <- counts %>% mutate(method = factor(method, levels = 
    c("gwas", "predixcan_sqtl", "predixcan_eqtl", "predixcan_sqtl_enloc", "predixcan_eqtl_enloc",
        "predixcan_sqtl_all_pairs", "predixcan_eqtl_all_pairs",  "predixcan_sqtl_enloc_all_pairs", "predixcan_eqtl_enloc_all_pairs",
        "predixcan_sqtl_mrcp", "predixcan_eqtl_mrcp",
        "multixcan_sqtl", "multixcan_eqtl", "multixcan_sqtl_enloc", "multixcan_eqtl_enloc")))
order_ <- counts %>% filter(method == "gwas") %>% arrange(n) %>% .$abbreviation
counts <- counts %>% mutate(abbreviation = factor(abbreviation, levels=order_))

predixcan_by_tissue_counts <- "data/gwas_hits/pred_by_tissue_count.txt" %>% r_tsv_
predixcan_by_tissue_counts <- predixcan_by_tissue_counts %>% mutate(abbreviation = factor(abbreviation, levels = order_))
```


```{r misc_functions, echo=FALSE}
theme_ <-function() {
  # theme_bw() +
  #   theme(plot.title = element_text(hjust=0.5, face="bold", size=27),
  #         plot.subtitle = element_text(hjust=0.5, face="italic", size=25),
  #         axis.title = element_text(size=25),
  #         axis.text = element_text(size=20),
  #         axis.text.x = element_text(size=15, angle=90, hjust=1),
  #         legend.text = element_text(size = 15),
  #         legend.position="bottom")
  theme_bw() + theme(legend.position="bottom")
}
```

The following figures compare the number of GWAS detections to predixcan and multixcan detections.

For every trait, the number of significant predixcan (gene,tissue) associations from expression models is proportional to the number of significant GWAS associations.

```{r predixcan_all_pairs, echo=FALSE}
  counts %>% filter(method == "gwas") %>% select(trait=trait, gwas=n) %>%
    inner_join(counts %>% filter(method !="gwas"), by="trait") %>%
    filter(method %in% c("predixcan_eqtl_all_pairs", "predixcan_eqtl_enloc_all_pairs", "predixcan_eqtl_mrcp")) %>% ggplot() + theme_() +
    geom_point(aes(x=gwas, y=n, color=category), size=4) + facet_wrap(~method, ncol=3, scales = "free_y") +
    scale_color_manual(values = palette_) + xlab("gwas detections") + ylab("predixcan detections") +
    ggtitle("Predixcan associations", subtitle = "All (gene,tissue) pairs")
```

This behavior is also observed when counting unique gene associations across all tissues:

```{r predixcan_genes, echo=FALSE}
  counts %>% filter(method == "gwas") %>% select(trait=trait, gwas=n) %>%
    inner_join(counts %>% filter(method !="gwas"), by="trait")  %>% filter(method %in% c("predixcan_eqtl", "predixcan_eqtl_enloc")) %>% ggplot() + theme_() +
    geom_point(aes(x=gwas, y=n, color=category), size=4) + facet_wrap(~method, ncol=2, scales = "free_y") +
    scale_color_manual(values = palette_) + xlab("gwas detections") + ylab("predixcan detections") +
    ggtitle("Predixcan associations", subtitle="unique genes")
```

And also with MultiXcan associations, able to detect more genes than single-tissue PrediXcan:

```{r multixcan_genes, echo=FALSE}
  counts %>% filter(method == "gwas") %>% select(trait=trait, gwas=n) %>%
    inner_join(counts %>% filter(method !="gwas"), by="trait") %>%
    filter(grepl("multixcan", method) & grepl("eqtl", method)) %>% ggplot() + theme_() +
    geom_point(aes(x=gwas, y=n, color=category), size=4) + facet_wrap(~method, ncol=2, scales = "free_y") +
    scale_color_manual(values = palette_) + xlab("gwas detections") + ylab("multixcan detections")
```


Examining associations within each tissue, we also observe a trend of increasing number of predixcan associations with number of GWAS associations.

```{r predixcan_single_tissue, echo=FALSE}

  counts %>% filter(method == "gwas") %>% select(trait=trait, gwas=n) %>%
    inner_join(predixcan_by_tissue_counts, by="trait") %>% 
    mutate(method = factor(method, levels = c("predixcan_sqtl", "predixcan_sqtl_enloc", "predixcan_eqtl", "predixcan_eqtl_enloc"))) %>%
    filter(grepl("predixcan_eqtl_enloc", method)) %>% mutate(gwas = ifelse(gwas>30000, 30000, gwas)) %>%  ggplot() + theme_() + 
    ggtitle("Predixcan associations", subtitle = "associations aggregated by tissue") +
    geom_boxplot(aes(gwas, n, group = cut_width(gwas, 1000))) + xlab("gwas associations") + ylab("Predixcan enlocalized associations")
``` 
