---
output:
  html_document: default
  pdf_document: default
  word_document: default
---

---
title: "Miscellaneous statistics for the GTEx-GWAS paper"
output:
  html_document: default
  pdf_document: default
---

#### Import libraries and source scripts
```{r import}
# install.packages("bigrquery")
suppressPackageStartupMessages(library(bigrquery))

# Script with ancillary functions to perform queries to BigQuery
suppressPackageStartupMessages(source("code/BigQuery.R", chdir = TRUE))
```

### Select phenotypes without deflation after imputation
```{r select_phenotypes}
NO_DEFLATION <- 0
STRONG_DEFLATION <- 1
MILD_DEFLATION <- 2

query <- glue::glue("SELECT Tag as phenotype, Deflation
                     FROM {gwas_metadata_tbl$dataset_name}.{gwas_metadata_tbl$table_name}")

gwas_deflation_status <- query_exec(query, gwas_metadata_tbl$project)

selected_phenotypes <- gwas_deflation_status %>% filter(Deflation == NO_DEFLATION) %>% .[,1]
strong_deflation_phenotypes <- gwas_deflation_status %>% filter(Deflation == STRONG_DEFLATION) %>% .[,1]
mild_deflation_phenotypes <- gwas_deflation_status %>% filter(Deflation == MILD_DEFLATION) %>% .[,1]

gwas_metadata <- gtex_gwas_metadata() %>% 
                 rename(phenotype="Tag") %>% 
                 select(phenotype, new_Phenotype, Category, Deflation)

gtex_tissue_metadata <- basicQuery(gtex_tissue_metadata_tbl) %>% 
                        select(tissue, v8_all, tissue_abbrv) %>% 
                        rename(sample_size=v8_all)
```
**`r length(selected_phenotypes)` phenotypes were selected for analysis. From the rest, `r length(strong_deflation_phenotypes)` GWAS showed strong deflation after imputation, whereas `r length(mild_deflation_phenotypes)` presented mild deflation.**


### PrediXcan and MultiXcan
```{r n_gene_mx, echo=FALSE}
query <- "SELECT 
  gene FROM (
  SELECT px.phenotype as phenotype, 
         px.gene as gene,
         px.tissue as tissue,
         gwas.Deflation as deflation 
  FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px 
  JOIN {gwas_metadata_tbl$dataset_name}.{gwas_metadata_tbl$table_name} as gwas 
  ON px.phenotype = gwas.Tag 
  WHERE gwas.Deflation = 0)
GROUP BY gene" %>% glue::glue()

n_genes <- query_exec(query, project = gwas_metadata_tbl$project, max_pages = Inf) %>% nrow

query <- "SELECT gene_type, COUNT(DISTINCT(gene)) as count 
FROM GTEx_V8_ElasticNet_EUR_2018_07_05.extra
GROUP BY gene_type"

genes_by_type <- query_exec(query, gwas_metadata_tbl$project)

rownames(genes_by_type) <- genes_by_type$gene_type
```
- **`r n_genes` genes are being tested by MultiXcan.**


```{r n_gene_tissue, echo=FALSE}
query <- "SELECT 
  gene, 
  tissue 
  FROM (
  SELECT px.phenotype as phenotype,
         px.gene as gene,
         px.tissue as tissue,
         gwas.Deflation as deflation 
  FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px 
  JOIN {gwas_metadata_tbl$dataset_name}.{gwas_metadata_tbl$table_name} as gwas 
  ON px.phenotype = gwas.Tag 
  WHERE gwas.Deflation = 0) 
GROUP BY gene, tissue" %>% glue::glue()

n_gene_tissue_pairs <- query_exec(query, project = gwas_metadata_tbl$project, max_pages = Inf) %>% nrow

pvalue_threshold <- pval_threshold_px <- 0.05 / n_gene_tissue_pairs
```
- **`r n_gene_tissue_pairs` (gene, tissue) pairs are being tested by PrediXcan.**


```{r gene_annot, echo=FALSE}
query <- glue::glue("SELECT gene_id, gene_type from annotations.gencode_v26")
gene_annot <- query_exec(query, project="gtex-awg-im", max_pages = Inf) %>% suppressWarnings()
```

```{r, echo=FALSE}
query <- "SELECT 
  best_mx.*,
  0.05/best_mx.n_genes as pvalue_thr,
  metadata.Deflation
FROM (
  SELECT phenotype, min(pvalue) as best_pvalue, count(*) as n_genes
  FROM {multixcan_tbl$dataset_name}.{multixcan_tbl$table_name}
  GROUP BY phenotype
) as best_mx
JOIN {gwas_metadata_tbl$dataset_name}.{gwas_metadata_tbl$table_name} metadata 
ON metadata.Tag = best_mx.phenotype" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf, use_legacy_sql = F) %>% suppressMessages()
str(df)
```
**`r sum(df$best_pvalue < df$pvalue_thr & df$Deflation == 0)` traits have at least one gene significantly associated (according to MultiXcan).**


```{r, echo=FALSE}
query <- "SELECT 
  best_px.*, 
  0.05/best_px.n_tissues_genes as pvalue_thr, 
  metadata.Deflation
FROM (
  SELECT phenotype, min(pvalue) as best_pvalue, count(*) as n_tissues_genes
  FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name}
  GROUP BY phenotype
) as best_px
JOIN {gwas_metadata_tbl$dataset_name}.{gwas_metadata_tbl$table_name} metadata 
ON metadata.Tag = best_px.phenotype" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf, use_legacy_sql = F) %>% suppressMessages()
str(df)

```
**`r sum(df$best_pvalue < df$pvalue_thr & df$Deflation == 0)` traits have at least one gene significantly associated in some tissue (according to PrediXcan).**


### PrediXcan and colocalization methods
#### How many among PrediXcan-significant are Coloc-colocalized?
```{r predixcan_coloc, echo=FALSE}
query <- "SELECT 
  px.phenotype as phenotype, 
  px.tissue as tissue, 
  px.gene as gene, 
  px.pvalue as pvalue, 
  coloc.PP_H4_abf as P4
FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px
JOIN {coloc_tbl$dataset_name}.{coloc_tbl$table_name} as coloc
ON px.phenotype = coloc.phenotype AND px.tissue = coloc.tissue AND px.gene = coloc.gene_id
WHERE pvalue < {pvalue_threshold}" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf) %>% suppressWarnings()
df <- df %>% filter(phenotype %in% selected_phenotypes)

str(df)

n_predixcan_signif <- df %>% group_by(gene) %>% count() %>% nrow
n_coloc <- df %>% filter(P4 > 0.5) %>% group_by(gene) %>% count() %>% nrow
```
**`r n_coloc` out of `r n_predixcan_signif` (`r round(100*(n_coloc/n_predixcan_signif), 2)`%)**


What if restricting to protein-coding genes?
```{r predixcan_coloc_protein_coding, echo=FALSE}
df <- df %>% mutate(gene = substr(gene, 1, 15))
df_protein_coding <- inner_join(df, gene_annot, by = c("gene"="gene_id")) %>% 
                     filter(gene_type == "protein_coding")
str(df_protein_coding)

n_predixcan_signif_pc <- df_protein_coding %>% group_by(gene) %>% count %>% nrow()
n_coloc_among_px_pc <- df_protein_coding %>% filter(P4 > 0.5) %>% group_by(gene) %>% count() %>% nrow()

```
**`r n_coloc_among_px_pc` out of `r n_predixcan_signif_pc` (`r round(100*n_coloc_among_px_pc / n_predixcan_signif_pc, 2)`%)**


#### How many among PrediXcan-significant are Enloc-colocalized?
```{r predixcan_enloc, echo=FALSE}
query <- "SELECT 
  px.phenotype as phenotype, 
  px.tissue as tissue, 
  px.gene as gene, 
  px.pvalue as pvalue, 
  enloc.rcp as rcp
FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px
JOIN {enloc_tbl$dataset_name}.{enloc_tbl$table_name} as enloc
ON px.phenotype = enloc.phenotype AND px.tissue = enloc.tissue AND px.gene = enloc.gene_id
WHERE pvalue < {pvalue_threshold}" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf) %>% suppressWarnings()
df <- df %>% filter(phenotype %in% selected_phenotypes)
str(df)

n_enloc <- df %>% filter(rcp > 0.5) %>% group_by(gene) %>% count() %>% nrow
```
**`r n_enloc` out of `r n_predixcan_signif` (`r round(100*(n_enloc/n_predixcan_signif), 2)`%)**

What if restricting to protein-coding genes?
```{r, echo=FALSE}
df <- df %>% mutate(gene = substr(gene, 1, 15))
df_protein_coding <- inner_join(df, gene_annot, by = c("gene"="gene_id")) %>% 
                     filter(gene_type == "protein_coding")
str(df_protein_coding)
n_enloc_among_px_pc <- df_protein_coding %>% filter(rcp > 0.5) %>% group_by(gene) %>% count() %>% nrow()
```
**`r n_enloc_among_px_pc` out of `r n_predixcan_signif_pc` (`r 100*(n_enloc_among_px_pc / n_predixcan_signif_pc)`%)**


#### How many among Coloc-colocalized genes are PrediXcan-significant?
```{r coloc_predixcan, echo=FALSE}
query <- "SELECT 
  px.phenotype as phenotype, 
  px.tissue as tissue, 
  px.gene as gene, 
  px.pvalue as pvalue, 
  coloc.PP_H4_abf as P4
FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px
JOIN {coloc_tbl$dataset_name}.{coloc_tbl$table_name} as coloc
ON px.phenotype = coloc.phenotype AND px.tissue = coloc.tissue AND px.gene = coloc.gene_id
WHERE PP_H4_abf > 0.5" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf) %>% suppressWarnings()
df <- df %>% filter(phenotype %in% selected_phenotypes)
str(df)

n_coloc <- df %>% group_by(gene) %>% count() %>% nrow()
n_predixcan <- inner_join(df, df %>% group_by(gene) %>% summarise(pthr=0.05/n()), by = "gene") %>% 
  filter(pvalue < pthr) %>% 
  group_by(gene) %>% 
  count() %>% 
  nrow()

```
**`r n_predixcan` out of `r n_coloc` (`r round(100*n_predixcan/n_coloc, 2)`%)**


#### How many among Enloc-colocalized genes are PrediXcan-significant?
```{r enloc_predixcan, echo=FALSE}
query <- "SELECT 
  px.phenotype as phenotype, 
  px.tissue as tissue, 
  px.gene as gene, 
  px.pvalue as pvalue, 
  enloc.rcp as rcp
FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px
JOIN {enloc_tbl$dataset_name}.{enloc_tbl$table_name} as enloc
ON px.phenotype = enloc.phenotype AND px.tissue = enloc.tissue AND px.gene = enloc.gene_id
WHERE rcp > 0.5" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf) %>% suppressWarnings()
df <- df %>% filter(phenotype %in% selected_phenotypes)
str(df)
n_enloc <- df %>% group_by(gene) %>% count() %>% nrow()
n_predixcan <- inner_join(df, df %>% group_by(gene) %>% summarise(pthr=0.05/n()), by = "gene") %>% 
  filter(pvalue < pthr) %>% 
  group_by(gene) %>% 
  count() %>% 
  nrow()
```
**`r n_predixcan` out of `r n_enloc` (`r round(100*n_predixcan/n_enloc, 2)`%)**


#### How many among Enloc-colocalized genes are Coloc-colocalized?
```{r enloc_coloc, echo=FALSE}
query <- "SELECT 
  coloc.phenotype as phenotype, 
  coloc.tissue as tissue, 
  coloc.gene_id as gene, 
  coloc.PP_H4_abf as P4, 
  enloc.rcp as rcp
FROM {coloc_tbl$dataset_name}.{coloc_tbl$table_name} as coloc
JOIN {enloc_tbl$dataset_name}.{enloc_tbl$table_name} as enloc
ON coloc.phenotype = enloc.phenotype AND coloc.tissue = enloc.tissue AND coloc.gene_id = enloc.gene_id
WHERE rcp > 0.5" %>% glue::glue()

df <- query_exec(query, project = predixcan_tbl$project, max_pages = Inf) %>% suppressWarnings()
df <- df %>% filter(phenotype %in% selected_phenotypes)
str(df)
n_enloc <- df %>% group_by(gene) %>% count() %>% nrow()
n_coloc <- df %>% filter(P4 > 0.5) %>% group_by(gene) %>% count() %>% nrow()
```
**`r n_coloc` out of `r n_enloc` (`r round(100*n_coloc/n_enloc, 2)`%)**


### How many loci with GWAS hits also contain a significant gene-level association?
```{r ldblock_counts}

# for the script to generate the file below, see ../code/count_gwas_vs_predixcan_hits.R
counts.per.ldblock <- readRDS("output/counts_per_ldblock_gwas_mx_and_enloc.rds")

# set levels so that phenotypes are displayed ordered according to the number of GWAS loci.
ordered_phenotypes <- counts.per.ldblock %>% arrange(gwas_counts) %>% select(phenotype, new_Phenotype, abbreviation)
counts.per.ldblock$phenotype <- factor(as.character(counts.per.ldblock$phenotype), levels = ordered_phenotypes$phenotype)
counts.per.ldblock$new_Phenotype <- factor(as.character(counts.per.ldblock$new_Phenotype), levels = ordered_phenotypes$new_Phenotype)
counts.per.ldblock$abbreviation <- factor(as.character(counts.per.ldblock$abbreviation), levels = ordered_phenotypes$abbreviation)

str(counts.per.ldblock)

cols <- c("GWAS"="#0000C0", "MultiXcan"="#800080", "MultiXcan + Enloc"="#00C000")
```


The following plots represent:

* The number of Pickrell regions with a GWAS hit (a proxy for the number of independent GWAS signals). The p-value threshold used was 0.05/`r as.integer(0.05/attr(counts.per.ldblock, "gwas_pval_thr"))`.
* Among those, the ones with a MultiXcan-significant gene. The p-value threshold used was 0.05/`r as.integer(0.05/attr(counts.per.ldblock, "mx_pval_thr"))`.
* The number of regions with a gene that is both MultiXcan-significant and Enloc-colocalized. The threshold for Enloc's RCP was `r attr(counts.per.ldblock, "rcp_thr")`


```{r ldblock_counts_fig1, fig.width=10, fig.asp=0.8, echo=FALSE}
GWAS_COUNT_THR <- 10 # discard traits with less than this number of GWAS hits
counts.per.ldblock_ <- counts.per.ldblock %>% filter(gwas_counts > GWAS_COUNT_THR)
pp <- ggplot() + 
  geom_point(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_counts, col="GWAS"), size=1.5) +
  geom_line(data = counts.per.ldblock_, mapping = aes(as.numeric(abbreviation)-min(as.numeric(abbreviation))+1, gwas_counts, col="GWAS"), group=1) +
  geom_bar(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_mx_counts, fill="MultiXcan"), stat="identity") +
  # geom_bar(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_mx_enloc_counts_region, fill="MultiXcan + Enloc (matched by region)"), stat="identity") +
  geom_bar(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_mx_enloc_counts_gene, fill="MultiXcan + Enloc"), stat="identity") +
  theme_bw() + theme(axis.text.x=element_text(angle=55, hjust = 1)) +
  xlab("Phenotype") + ylab("# of LD-independent blocks with a significant association") + 
  scale_y_continuous(trans="log10") +
  scale_fill_manual(name="Method", values=cols) +
  scale_color_manual(label="", values=cols)

pp
```


The same, but by trait category:
```{r ldblock_counts_fig2, fig.width=10, fig.asp=1.5, echo=FALSE}
pp <- ggplot() + 
  geom_point(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_counts, col="GWAS"), size=1.5) + 
  geom_bar(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_mx_counts, fill="MultiXcan"), stat="identity") + 
  # geom_bar(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_mx_enloc_counts_region, fill="MultiXcan + Enloc (matched by region)"), stat="identity") + 
  geom_bar(data = counts.per.ldblock_, mapping = aes(x = new_Phenotype, y = gwas_mx_enloc_counts_gene, fill="MultiXcan + Enloc"), stat="identity") + 
  theme_bw() + theme(axis.text.x=element_text(angle=55, hjust = 1)) + 
  xlab("Phenotype") + ylab("# of LD-independent blocks with a significant association") + 
  scale_y_continuous(trans="log10") + 
  scale_fill_manual(name="Method", values=cols) + 
  scale_color_manual(label="", values=cols) + 
  facet_wrap(Category ~ ., drop = T, scales="free_x", nrow = 3)

pp
```


### Do primary eQTLs have larger effect on complex traits than secondary eQTLs?
*t*-tests were performed for each (tissue, trait) pair:
```{r ttest_prim_vs_sec_, echo=FALSE}
# for the script to generate the file below, see t_test_primary_vs_secondary.R
t.test_df <- readRDS("output/t_test_primary_vs_secondary.rds")

df <- inner_join(t.test_df, gtex_tissue_metadata, by="tissue") %>%
      inner_join(gwas_metadata, by="phenotype") %>% suppressWarnings()

df$tissue <- factor(df$tissue, levels = gtex_tissue_metadata[order(gtex_tissue_metadata$sample_size), "tissue"])

phenotype <- df %>% group_by(phenotype) %>% summarise(median(t.test_pvalue))
df$phenotype <- factor(df$phenotype, levels = phenotype[order(-phenotype$`median(t.test_pvalue)`), "phenotype"]$phenotype)
phenotype <- df %>% group_by(new_Phenotype) %>% summarise(median(t.test_pvalue))
df$new_Phenotype <- factor(df$new_Phenotype, levels = phenotype[order(-phenotype$`median(t.test_pvalue)`), "new_Phenotype"]$new_Phenotype)
```


```{r ttest_prim_vs_sec_fig_1, fig.width = 10, fig.asp = 0.7}
ggp1 <- ggplot(df, aes(x=tissue, y=-log10(t.test_pvalue))) + 
        geom_boxplot() + 
        theme_bw() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
ggp1
```


```{r ttest_prim_vs_sec_fig_2, fig.width = 10, fig.asp = 1.8}
ggp2 <- ggp1 + facet_grid(Category ~ .)
ggp2
```


```{r ttest_prim_vs_sec_fig_3, fig.width = 10, fig.asp = 0.7}
  # ggp2 <- ggplot(df, aes(x=phenotype, y=-log10(t.test_pvalue))) + geom_boxplot() + theme_bw() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
ggp3 <- ggplot(df, aes(x=new_Phenotype, y=-log10(t.test_pvalue))) +
        geom_boxplot() +
        theme_bw() + theme(axis.text.x = element_text(angle = 60, hjust = 1))
ggp3
```


```{r ttest_prim_vs_sec_fig_4, fig.width = 10, fig.asp = 1.5}
ggp4 <- ggp3 + facet_wrap(Category ~ ., drop = T, scales="free_x", nrow = 3)
ggp4
```


### How does variance of gene effects vary across traits?
```{r variance_beta_gene}
beta_gene_variance_df <- readRDS("output/beta_gene_variance.rds")
beta_gene_variance_df <- beta_gene_variance_df %>% mutate(beta_gene_var_norm=var_beta_gene/var_delta)
phenotype <- beta_gene_variance_df %>% group_by(phenotype) %>% summarise(median(var_beta_gene)) %>% inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")
beta_gene_variance_df <- beta_gene_variance_df %>% inner_join(gwas_metadata, by="phenotype") %>% inner_join(gtex_tissue_metadata, by="tissue") %>% rename(n_samples_gtex=sample_size)
beta_gene_variance_df$phenotype <- factor(beta_gene_variance_df$phenotype, levels = phenotype[order(-phenotype$`median(var_beta_gene)`), "phenotype"]$phenotype)
beta_gene_variance_df$new_Phenotype <- factor(beta_gene_variance_df$new_Phenotype, levels = phenotype[order(-phenotype$`median(var_beta_gene)`), "new_Phenotype"]$new_Phenotype)

ggp <- ggplot(beta_gene_variance_df, aes(x=new_Phenotype, y=var_beta_gene))
ggp <- ggp + geom_boxplot() + theme_bw()
ggp <- ggp + theme(axis.text.x=element_text(angle=65, hjust = 1))

```

```{r variance_beta_gene_fig1, fig.width = 10, fig.asp = 0.7}
ggp
```

```{r variance_beta_gene_fig2, fig.width = 10, fig.asp = 1.8}
ggp <- ggp + facet_wrap(Category ~ ., drop = T, scales="free_x", nrow = 3)
ggp
```

```{r}
knitr::kable(beta_gene_variance_df %>% select(phenotype, tissue, var_beta_gene) %>% group_by(phenotype) %>% top_n(3, wt=var_beta_gene) %>% dplyr::arrange(phenotype, -var_beta_gene))
```

Some observations:

* The tissue with largest $\text{Var}(\beta_{\text{gene}})$ for the blood cell count traits is whole blood (except for red blood cells where it is heart left ventricle).
* The tissue with largest $\text{Var}(\beta_{\text{gene}})$ for rheumatoid arthritis is liver. There is evidence of a [connection between RA and liver damage](https://www.ncbi.nlm.nih.gov/pubmed/29627896).
* The values for height from GIANT are not reproduced in UKB.


### Fraction of variance in GWAS effect size explained by variance in eQTL effect size
$$\delta_{\text{GWAS}}=\delta_0+\alpha\times \text{LDScore} + \beta_{\text{gene}}\times \gamma_{\text{eQTL}}+\epsilon$$

$$\text{Var}(\delta_{\text{GWAS}}-\alpha\times \text{LDScore}-\delta_0)=\text{Var}(\delta')=\underbrace{\text{Var}(\beta)\text{Var}(\gamma_{\text{eQTL}})}_{=:V}+\text{Var}(\epsilon)$$
$$r:=\frac{V}{\text{Var}(\delta')}=1-\frac{\text{Var}(\epsilon)}{\text{Var}(\delta')}$$

```{r fraction_explained}
beta_gene_variance_df_ <- readRDS("output/beta_gene_variance.rds")

phenotype <- beta_gene_variance_df_ %>% 
             group_by(phenotype) %>%
             summarise(median(var_beta_gene)) %>%
             inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")

beta_gene_variance_df_ <- beta_gene_variance_df_ %>%
                          inner_join(gwas_metadata, by="phenotype") %>%
                          inner_join(gtex_tissue_metadata, by="tissue") %>%
                          rename(n_samples_gtex=sample_size)

beta_gene_variance_df_ <- beta_gene_variance_df_ %>% filter(n_samples_gtex > 400)
phenotype <- beta_gene_variance_df_ %>% group_by(phenotype) %>% summarise(median(fraction_explained)) %>% inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")
beta_gene_variance_df_$phenotype <- factor(beta_gene_variance_df_$phenotype, levels=phenotype[order(-phenotype$`median(fraction_explained)`), "phenotype"]$phenotype)
beta_gene_variance_df_$new_Phenotype <- factor(beta_gene_variance_df_$new_Phenotype, levels=phenotype[order(-phenotype$`median(fraction_explained)`), "new_Phenotype"]$new_Phenotype)
```

```{r, fig.width = 10, fig.asp = 0.7}
ggp <- ggplot(beta_gene_variance_df_, aes(x=new_Phenotype, y=fraction_explained))
ggp <- ggp + geom_boxplot() + theme_bw()
ggp <- ggp + theme(axis.text.x=element_text(angle=65, hjust = 1))
ggp
```

```{r, fig.width = 10, fig.asp = 1.8}
ggp <- ggp + facet_wrap(Category ~ ., drop = T, scales="free_x", nrow = 3)
ggp
```


We note that comparisons across tissues within a single trait are not meaningful because variance in $\gamma$ is overestimated for tissues with lower sample sizes.
Fraction of $\delta'$ explained by expression is thus overestimated as well in these tissues.

```{r}
ggp <- ggplot(beta_gene_variance_df, aes(x=n_samples_gtex, y=fraction_explained)) + geom_smooth() + theme_bw()
ggp
```


```{r}
ggp <- ggplot(beta_gene_variance_df %>%
              group_by(tissue) %>% 
              summarise(mean_var_gamma=mean(var_gamma)) %>%
              inner_join(gtex_tissue_metadata %>% select("tissue", "sample_size") %>% rename(n_samples_gtex=sample_size), by="tissue"),
              aes(x=n_samples_gtex, y=mean_var_gamma)) +
       geom_point() + theme_bw()
ggp
```


```{r}
knitr::kable(beta_gene_variance_df_ %>% select(phenotype, tissue, fraction_explained) %>% group_by(phenotype) %>% top_n(1, wt=fraction_explained))
```

```{r}
summary_dir <- "output/lmm/lme4_model_2"
beta_gene_variance_df_ <- lapply(file.path(summary_dir, list.files(summary_dir)), function(x) read.table(x, header=TRUE, stringsAsFactors = FALSE))
beta_gene_variance_df_ <- bind_rows(beta_gene_variance_df_) %>% suppressWarnings()

beta_gene_variance_df_ <- beta_gene_variance_df_ %>%
                          inner_join(gwas_metadata, by="phenotype") %>%
                          inner_join(gtex_tissue_metadata, by="tissue") %>%
                          rename(n_samples_gtex=sample_size)

phenotype <- beta_gene_variance_df_ %>% group_by(phenotype) %>% summarise(median(var_beta_gene)) %>% inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")
beta_gene_variance_df_$phenotype <- factor(beta_gene_variance_df_$phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "phenotype"]$phenotype)
beta_gene_variance_df_$new_Phenotype <- factor(beta_gene_variance_df_$new_Phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "new_Phenotype"]$new_Phenotype)

```

```{r, fig.width = 10, fig.asp = 0.7}
ggp <- ggplot(beta_gene_variance_df_, aes(x=new_Phenotype, y=var_beta_gene))
ggp <- ggp + geom_boxplot() + theme_bw()
ggp <- ggp + theme(axis.text.x=element_text(angle=65, hjust = 1))
ggp <- ggp + xlab("phenotype") + ylab("fraction of beta_gwas variance explained by beta_eqtl variance")
ggp
```


```{r}
summary_dir <- "output/lmm/lme4_model_3"
beta_gene_variance_df_ <- lapply(file.path(summary_dir, list.files(summary_dir)), function(x) read.table(x, header=TRUE, stringsAsFactors = FALSE))
beta_gene_variance_df_ <- bind_rows(beta_gene_variance_df_) %>% suppressWarnings()

beta_gene_variance_df_ <- beta_gene_variance_df_ %>%
                          inner_join(gwas_metadata, by="phenotype") %>%
                          inner_join(gtex_tissue_metadata, by="tissue") %>%
                          rename(n_samples_gtex=sample_size)

phenotype <- beta_gene_variance_df_ %>% group_by(phenotype) %>% summarise(median(var_beta_gene)) %>% inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")
beta_gene_variance_df_$phenotype <- factor(beta_gene_variance_df_$phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "phenotype"]$phenotype)
beta_gene_variance_df_$new_Phenotype <- factor(beta_gene_variance_df_$new_Phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "new_Phenotype"]$new_Phenotype)

```

```{r, fig.width = 10, fig.asp = 0.7}
ggp <- ggplot(beta_gene_variance_df_, aes(x=new_Phenotype, y=var_beta_gene))
ggp <- ggp + geom_boxplot() + theme_bw()
ggp <- ggp + theme(axis.text.x=element_text(angle=65, hjust = 1))
ggp <- ggp + xlab("phenotype") + ylab("var_beta_gene")
ggp
```


MASHR with effect sizes
```{r}
summary_dir <- "output/lmm/lme4_model_6"
beta_gene_variance_df_ <- lapply(file.path(summary_dir, list.files(summary_dir)), function(x) read.table(x, header=TRUE, stringsAsFactors = FALSE))
beta_gene_variance_df_ <- bind_rows(beta_gene_variance_df_) %>% suppressWarnings()

beta_gene_variance_df_ <- beta_gene_variance_df_ %>%
                          inner_join(gwas_metadata, by="phenotype") %>%
                          inner_join(gtex_tissue_metadata, by="tissue") %>%
                          rename(n_samples_gtex=sample_size)

phenotype <- beta_gene_variance_df_ %>% group_by(phenotype) %>% summarise(median(var_beta_gene)) %>% inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")
beta_gene_variance_df_$phenotype <- factor(beta_gene_variance_df_$phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "phenotype"]$phenotype)
beta_gene_variance_df_$new_Phenotype <- factor(beta_gene_variance_df_$new_Phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "new_Phenotype"]$new_Phenotype)

```

```{r, fig.width = 10, fig.asp = 0.7}
ggp <- ggplot(beta_gene_variance_df_, aes(x=new_Phenotype, y=var_beta_gene))
ggp <- ggp + geom_boxplot() + theme_bw()
ggp <- ggp + theme(axis.text.x=element_text(angle=65, hjust = 1))
ggp <- ggp + xlab("phenotype") + ylab("var_beta_gene")
ggp
```


MASHR with zscores
```{r}
summary_dir <- "output/lmm/lme4_model_5"
beta_gene_variance_df_ <- lapply(file.path(summary_dir, list.files(summary_dir)), function(x) read.table(x, header=TRUE, stringsAsFactors = FALSE))
beta_gene_variance_df_ <- bind_rows(beta_gene_variance_df_) %>% suppressWarnings()

beta_gene_variance_df_ <- beta_gene_variance_df_ %>%
                          inner_join(gwas_metadata, by="phenotype") %>%
                          inner_join(gtex_tissue_metadata, by="tissue") %>%
                          rename(n_samples_gtex=sample_size)

phenotype <- beta_gene_variance_df_ %>% group_by(phenotype) %>% summarise(median(var_beta_gene)) %>% inner_join(gwas_metadata %>% select(-Deflation), by="phenotype")
beta_gene_variance_df_$phenotype <- factor(beta_gene_variance_df_$phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "phenotype"]$phenotype)
beta_gene_variance_df_$new_Phenotype <- factor(beta_gene_variance_df_$new_Phenotype, levels=phenotype[order(-phenotype$`median(var_beta_gene)`), "new_Phenotype"]$new_Phenotype)

```

```{r, fig.width = 10, fig.asp = 0.7}
ggp <- ggplot(beta_gene_variance_df_, aes(x=new_Phenotype, y=var_beta_gene))
ggp <- ggp + geom_boxplot() + theme_bw()
ggp <- ggp + theme(axis.text.x=element_text(angle=65, hjust = 1))
ggp <- ggp + xlab("phenotype") + ylab("var_beta_gene")
ggp
```
<!---
query <- glue::glue("SELECT  FROM {predixcan_tbl$dataset_name}.{predixcan_tbl$table_name} as px 
                     WHERE pvalue < {pval_threshold_px}")
df <- query_exec(query, project=predixcan_tbl$project, max_pages = Inf)
df <- df %>% filter(phenotype %in% selected_phenotypes)
-->