---
title: "miscellaneous_statistics_2"
author: "Alvaro"
date: "2019-08-20"
output: workflowr::wflow_html
---

## Introduction

Here follow miscellaneous numbers to be plugged in the main GTEx paper, supplements, and GWAs companion (and its supplement)

```{r, echo=FALSE, warning=FALSE, message=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(bigrquery))


suppressWarnings(source("code/helpers/_helpers.R"))
suppressWarnings(source("code/helpers/_helpers_big_query.R"))
suppressWarnings(source("code/helpers/_helpers_big_query_tables.R"))

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
gwas_metadata <- (function(){
  query <- glue::glue("SELECT Tag as phenotype, Deflation as deflation, new_abbreviation as abbreviation
                       FROM {gwas_metadata_tbl$dataset_name}.{gwas_metadata_tbl$table_name}
                       WHERE Deflation=0")
  query_exec(query, project = "gtex-awg-im", use_legacy_sql = FALSE, max_pages = Inf)
})()
pheno_whitelist <- sprintf("(%s)", toString(sprintf("'%s'", gwas_metadata$phenotype)))
```

# ENLOC runs

```{r, echo=FALSE, warning=FALSE, message=FALSE}
n_enloc_eqtl_eur_tuples <- (function(){
  query <- 
"SELECT COUNT(*) 
 FROM (SELECT DISTINCT molecular_qtl_trait as gene_id, phenotype, tissue
       FROM {enloc_tbl_eqtl_eur$dataset_name}.{enloc_tbl_eqtl_eur$table_name} as e
       WHERE phenotype in {pheno_whitelist})" %>% glue::glue()
 query_exec(query, project = "gtex-awg-im", use_legacy_sql = FALSE, max_pages = Inf)
})() %>% as.double
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
n_enloc_eqtl_eur_gt_pairs <- (function(){
  query <- "
SELECT COUNT(*)
FROM (
  SELECT DISTINCT gene_id, tissue
  FROM (
    SELECT molecular_qtl_trait as gene_id, phenotype, tissue
    FROM {enloc_tbl_eqtl_eur$dataset_name}.{enloc_tbl_eqtl_eur$table_name} as e
    WHERE phenotype in {pheno_whitelist}
  )
)" %>% glue::glue()
 query_exec(query, project = "gtex-awg-im", use_legacy_sql = FALSE, max_pages = Inf)
})() %>% as.double
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
n_enloc_sqtl_eur_tuples <- (function(){
  query <- 
"SELECT COUNT(*) 
 FROM (SELECT DISTINCT molecular_qtl_trait as intron_id, phenotype, tissue
       FROM {enloc_tbl_sqtl_eur$dataset_name}.{enloc_tbl_sqtl_eur$table_name} as e
       WHERE phenotype in {pheno_whitelist})" %>% glue::glue()
 query_exec(query, project = "gtex-awg-im", use_legacy_sql = FALSE, max_pages = Inf)
})() %>% as.double
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
n_enloc_sqtl_eur_gt_pairs <- (function(){
  query <- "
SELECT COUNT(*)
FROM (
  SELECT DISTINCT intron_id, tissue
  FROM (
    SELECT molecular_qtl_trait as intron_id, phenotype, tissue
    FROM {enloc_tbl_sqtl_eur$dataset_name}.{enloc_tbl_sqtl_eur$table_name} as e
    WHERE phenotype in {pheno_whitelist}
  )
)" %>% glue::glue()
 query_exec(query, project = "gtex-awg-im", use_legacy_sql = FALSE, max_pages = Inf)
})() %>% as.double
```

We obtained:

- `r n_enloc_eqtl_eur_tuples` `(gene, tissue, phenotype)` distinct tuples from expression on Eur individuals
- `r n_enloc_eqtl_eur_gt_pairs` `(gene, tissue)` distinct pairs from expression on Eur individuals
- `r n_enloc_sqtl_eur_tuples` `(gene, tissue, phenotype)` distinct tuples from splicing on Eur individuals
- `r n_enloc_sqtl_eur_gt_pairs` `(gene, tissue)` distinct pairs from splicing on Eur individuals

